---
title: "Class Activity 1"
author: "Your name here"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = FALSE) # document will compile even in case of error
```


The `R` package `babynames` provides data about the popularity of individual baby names from the US Social Security Administration. Data includes all names used at least 5 times in a year beginning in 1880.

```{r}
#install.packages("babynames")  # uncomment to install
library(babynames)
```

Below is the list for first few cases of baby names.

```{r}
head(babynames)
```

1. How many cases and variables are in the dataset `babynames`?

**Answer:** 

```{r}
dim(babynames)
```

There are `r dim(babynames)[1]` cases and `r dim(babynames)[2]` variables in the dataset `babynames`.

Let's use the package tidyverse to do some exploratory data analysis.

```{r}
#install.packages("tidyverse")   # uncomment to install
library(tidyverse)
babynames %>% filter(name=='Aimee')

filtered_names <- babynames %>% filter(name=='Aimee')
```

```{r}
#install.packages("ggplot2")   # uncomment to install
library(ggplot2)
```

```{r trend, fig.cap= "A trend chart", tidy=FALSE}
ggplot(data=filtered_names, aes(x=year, y=prop)) + 
  geom_line(aes(colour=sex)) + 
  xlab('Year') + 
  ylab('Prop. of Babies Named Aimee')
```

2. What do you see in the Figure 1? Explain in a few sentences.

**Answer:**

\vspace*{0.5in}


3. Repeat question 2 to infer how does the proportion of babies with your first name trend over time. Examine the generated plot and describe the trend of your name's popularity over time. Consider the following points:

Has the popularity of your name increased, decreased, or remained stable over the years? Is there a noticeable difference in popularity between sexes? Are there any interesting patterns or trends, such as sudden increases or decreases in popularity?

**Answer:**

\vspace*{0.5in}


```{r}
# Replace 'YourName' with your first name
your_name <- "Dee"

your_name_data <- babynames %>% filter(name == your_name)

ggplot(data=your_name_data, aes(x=year, y=prop)) + 
  geom_line(aes(colour=sex)) + 
  xlab('Year') + 
  ylab(paste('Prop. of Babies Named', your_name))
```

4. Compare the popularity of your first name with a randomly chosen name from the dataset. Examine the generated plot and compare the popularity of your first name with the randomly chosen name. Consider the following points:

Are there differences in popularity trends between the two names? Is one name consistently more popular than the other, or do their popularity levels change over time? Are there any interesting patterns or trends in the data, such as periods of rapid increase or decrease in popularity?

**Answer**

\vspace*{0.5in}


```{r}
# Replace 'YourName' with your first name
your_name_data <- babynames %>% filter(name == 'Dee')

# Replace 'RandomName' with a randomly chosen name from the dataset
random_name_data <- babynames %>% filter(name == 'Max')

# Combine the two datasets
combined_data <- bind_rows(your_name_data, random_name_data)

# Plot the data
ggplot(data=combined_data, aes(x=year, y=prop)) + 
  geom_line(aes(colour=sex, linetype=name)) + 
  xlab('Year') 
```


# Extras (optional)

## Part 1: Setting Working Directory and Loading Data

1. Set your working directory to a folder on your computer where you would like to save your R scripts and data files.

```{r}
# Replace 'your_directory_path' with the path to your desired folder
# setwd("your_directory_path")
```


2. Load the mtcars dataset which comes preloaded with R. This dataset consists of various car features and their corresponding miles per gallon (mpg) values.


```{r}
data(mtcars)
head(mtcars)
```


## Part 2: Downloading Packages

1. Install the "tidyverse" package, which is a collection of useful R packages for data manipulation, exploration, and visualization.

```{r}
# Uncomment the line below to install the package
# install.packages("tidyverse")
```


2. Load the "tidyverse" package into your R session.


```{r}
library(tidyverse)
```


## Part 3: Creating and Compiling an R Markdown File


1. Create a new R Markdown file in RStudio by clicking on "File" > "New File" > "R Markdown...". Save the file in your working directory.

2. Add the following code to your R Markdown file to create a scatter plot of the mtcars dataset, showing the relationship between miles per gallon (mpg) and the weight of the car (wt).


```{r}
# Create a scatter plot
ggplot(data = mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  xlab("Weight (1000 lbs)") +
  ylab("Miles per Gallon") +
  theme_minimal()
```


3. Knit your R Markdown file to create an output document. Click the "Knit" button at the top of the RStudio script editor, and choose the output format you prefer (e.g., HTML, PDF, or Word).


\vspace*{0.2in}


## Questions

### 1. How does the weight of a car (wt) affect its miles per gallon (mpg) based on the scatter plot you created?

**Answer:**



\vspace*{0.2in}

### 2. What is the importance of setting a working directory in R?

**Answer:**





\vspace*{0.2in}

### 3. Explain the role of R Markdown in creating reproducible research documents.

**Answer:**



